---
title: "Lead IQ Report"
author: "Siyang Ren"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This report provides an analysis of the dataset "lead-iq-01.csv," which contains information about IQ levels of children aged 3 to 15 in El Paso, who lived at various distances from a large lead-emitting ore smelter. Lead exposure is known to affect neurological function, and this dataset captures IQ scores of children based on their proximity to the smelter. The dataset includes the columns:

- **Smelter**: Categorical variable indicating whether the subject lived "Near" or "Far" from the lead smelter.
- **IQ**: Numerical variable representing the IQ score as measured by the Wechsler Intelligence Scale for Children (WISC).

We will explore the data using visualizations and descriptive statistics.

## Load Data

```{r load-data, echo = FALSE, message = FALSE}
# Load required libraries
library(ggplot2)
library(dplyr)
library(knitr)

# Load the data
lead_data <- read.csv("../data-raw/lead-iq-01.csv")
```

## Graph: IQ Levels by Location Status

The graph below shows the distribution of IQ levels between the two locations, "Near" and "Far". Note that an outlier with an IQ value of "999" has been excluded from this figure:

```{r plot-iq}
ggplot(lead_data[lead_data$IQ < 999, ], aes(x = Smelter, y = IQ, fill = Smelter)) +
  geom_boxplot() +
  labs(title = "IQ Levels by Location Status",
       x = "Location",
       y = "IQ Level") +
  theme_minimal()
```

## Summary Table


```{r summary-table}
# Calculate summary statistics
summary_table <- lead_data %>%
  group_by(Smelter) %>%
  summarise(
    Mean_IQ = round(mean(IQ, na.rm = TRUE), 1),
    SD_IQ = round(sd(IQ, na.rm = TRUE), 1),
    Count = n()
  )

# Display the summary table
kable(summary_table, caption = "Summary Statistics for IQ Levels by Location Status")
```

The table above provides a summary of IQ levels by location status, including the mean and standard deviation for each group. The sample sizes for both categories are similar. The average IQ for the "Near" group is `r summary_table$Mean_IQ[summary_table$Smelter == "Near"]`, while the average IQ for the "Far" group is `r summary_table$Mean_IQ[summary_table$Smelter == "Far"]`. The mean IQ of children living far from the smelter appears noticeably higher compared to those living near the smelter. However, it is important to note that there is an outlier in the "Far" group, which makes it challenging to directly compare the standard deviations between the two groups.
